I can see clearly now: reinterpreting statistical significance

### 2018-10-10 doc.tex
Some hacky (meant to be quick) changes in formatting for submission to ecology

\begin{abstract}

Null hypothesis significance testing remains popular despite decades of concern about misuse and misinterpretation. We believe that much of the problem is due to language: significance testing has little to do with other meanings of the word ``significance''. Despite the limitations of null-hypothesis tests, we argue here that they remain useful in many contexts as a guide to whether a certain effect can be seen \emph{clearly} in that context (e.g. whether we can clearly see that a correlation or between-group difference is positive or negative). We therefore suggest that researchers describe the conclusions of null-hypothesis tests in terms of statistical ``clarity'' rather than statistical ``significance''. This simple semantic change could substantially enhance clarity in statistical communication.

\end{abstract}

@@

\section*{Introduction}

Statisticians and scientists have bemoaned the shortcomings of null hypothesis significance testing (NHST) for nearly a century \citep{Cohen1994}.
Books and articles proposing the de-emphasis or abandonment of the \pval have been cited thousands of times \citep{Cohen1994, Goodman1999, Wilkinson1999, ZiliakandMcCloskey2008, WassersteinandLazar2016}. These works plead for a focus on effect sizes and confidence intervals, and point out that null effects that truly have zero magnitude are unrealistic or impossible in most fields outside of the hard physical sciences \citep{Meehl1990, Tukey1991, Cohen1994}. Yet, \pvals without confidence intervals (or even effect sizes) and references to null effects still pervade the scientific literature at all levels up to and including articles in high-impact journals.

In a meta-analysis of 356 studies \citet{Bernardietal.2017} found that 72\% of studies contained an ambiguous use of the term ``significant'', 49\% interpreted non-significant effects as zero effects, and 44\% failed to report a comprehensible effect size. The misuse and misinterpretation of NHST is so frequent that there have been recent calls for drastically reducing \citep{SzucsandIoannidis2017} or abandoning \citep{McShaneetal.2017} its use. Other prescriptions have included the complete abandonment of frequentist statistics \citep{The2011}, or the use of a stricter significance threshold (e.g. $p < 0.005$: \citealt{Benjaminetal.2018}); however, the former seems impractical, while the latter is unlikely to reduce the misuse and misinterpretation of \pvals, or the publication bias imposed by any \pval threshold \citep{Ridleyetal.2007}.

@@

\section*{The null hypothesis is false}

In most biological studies, the null hypothesis is known or believed to not be strictly true.
Even in cases where the null hypothesis is sensible (e.g., particle physics, \cite{Staley2017}), NHST does not provide evidence that a difference is exactly zero.
This being the case, it is worth asking how NHST has survived ``if it is as idiotic as \ldots long believed'' \citet[cited in \cite{Kramer2011}]{ZiliakandMcCloskey2008}.

@@

\section*{The \pval is a property of the study}

We often see sentences like, ``X et al. showed that there is no significant effect of Y on Z'' with the implication that this effect can now be assumed to be absent (or unimportant). In fact, the sentence is erroneous even before we get to the implication: significance tests provide information about \emph{a data set} -- that is, about a study, not about the study system \citep{HoenigandHeisey2001}.
Indeed, a very small effect can lead to $p < 0.05$, when data is abundant (or noise is small); or a very large one can lead to $p > 0.05$ when the sample is small or noisy.

@@

\section*{Statistical Clarity}

The language of ``statistical clarity'' could help researchers escape various logical traps while interpreting the results of NHST, allowing for the continued use of NHST as a simple, robust method of evaluating whether a data signal is clear (see \citealt{Abelson1997} for arguments for NHST).
The use of ``significance'' to describe the results of hypothesis tests is deeply, and sometimes subtly, misleading, because it is at odds with other meanings of the word: the \pval is not an accurate gauge of whether a result is large in magnitude, biologically important, or relevant. ``Clarity,'' on the other hand, is an apt term for what NHST actually evaluates. 
\citet{jones2000sensible} and \citet{robinson2001past} suggest that researchers should report $p > 0.05$ using language such as ``the direction of the differences among the treatments was undetermined''. This is a step in the right direction. 
Replacing ``significance'' with ``clarity'' takes this idea further, and has the promise to substantially improve statistical communication.

@@

\input{table.tex}

@@

We believe that NHST is useful as a simple, robust way to ask whether an effect can be seen clearly in a particular data set \citep{robinson2001past}, and that careful, clarity-based language can reduce misinterpretation and miscommunication.

@@

We thank members of the Dushoff and Bolker labs for helpful comments on the first draft of the manuscript.

>>>>

@@

\linenumbers
\doublespacing

\subsection*{Abstract}

\noindent Null hypothesis significance testing remains popular despite decades of concern about misuse and misinterpretation. We believe that much of the problem is due to language: significance testing has little to do with other meanings of the word ``significance''. Despite the limitations of null-hypothesis tests, we argue here that they remain useful in many contexts as a guide to whether a certain effect can be seen \emph{clearly} in that context (e.g. whether we can clearly see that a correlation or between-group difference is positive or negative). We therefore suggest that researchers describe the conclusions of null-hypothesis tests in terms of statistical ``clarity'' rather than statistical ``significance''. This simple semantic change could substantially enhance clarity in statistical communication.

\subsection*{Key Words}

\noindent Statistical philosophy; Statistical clarity; Hypothesis testing; \pval

\subsection*{Introduction}

\noindent Statisticians and scientists have bemoaned the shortcomings of null hypothesis significance testing (NHST) for nearly a century \citep{Cohen1994}. Books and articles proposing the de-emphasis or abandonment of the \pval have been cited thousands of times \citep{Cohen1994, Goodman1999, Wilkinson1999, ZiliakandMcCloskey2008, WassersteinandLazar2016}. These works plead for a focus on effect sizes and confidence intervals, and point out that null effects that truly have zero magnitude are unrealistic or impossible in most fields outside of the hard physical sciences \citep{Meehl1990, Tukey1991, Cohen1994}. Yet, \pvals without confidence intervals (or even effect sizes) and references to null effects still pervade the scientific literature at all levels up to and including articles in high-impact journals.

In a meta-analysis of 356 studies \citet{Bernardietal.2017} found that 72\% of studies contained an ambiguous use of the term ``significant'', 49\% interpreted non-significant effects as zero effects, and 44\% failed to report a comprehensible effect size. The misuse and misinterpretation of NHST is so frequent that there have been recent calls for drastically reducing \citep{SzucsandIoannidis2017} or abandoning \citep{McShaneetal.2017} its use. Other prescriptions have included the complete abandonment of frequentist statistics \citep{The2011}, or the use of a stricter significance threshold (e.g. $p < 0.005$: \cite{Benjaminetal.2018}); however, the former seems impractical, while the latter is unlikely to reduce the misuse and misinterpretation of \pvals, or the publication bias imposed by any \pval threshold \citep{Ridleyetal.2007}.

@@

\subsection*{The null hypothesis is false}

\noindent In most biological studies, the null hypothesis is known or believed to not be strictly true. Even in cases where the null hypothesis is sensible (e.g., particle physics, \cite{Staley2017}), NHST does not provide evidence that a difference is exactly zero. This being the case, it is worth asking how NHST has survived ``if it is as idiotic as \ldots long believed'' \citet[cited in \cite{Kramer2011}]{ZiliakandMcCloskey2008}.

@@

\subsection*{The \pval is a property of the study}

\noindent We often see sentences like, ``X et al. showed that there is no significant effect of Y on Z'' with the implication that this effect can now be assumed to be absent (or unimportant). In fact, the sentence is erroneous even before we get to the implication: significance tests provide information about \emph{a data set} -- that is, about a study, not about the study system \citep{HoenigandHeisey2001}. Indeed, a very small effect can lead to $p < 0.05$, when data is abundant (or noise is small); or a very large one can lead to $p > 0.05$ when the sample is small or noisy.

@@

\subsection*{Statistical clarity}

\noindent The language of ``statistical clarity'' could help researchers escape various logical traps while interpreting the results of NHST, allowing for the continued use of NHST as a simple, robust method of evaluating whether a data signal is clear (see \cite{Abelson1997} for arguments for NHST). The use of ``significance'' to describe the results of hypothesis tests is deeply, and sometimes subtly, misleading, because it is at odds with other meanings of the word: the \pval is not an accurate gauge of whether a result is large in magnitude, biologically important, or relevant. ``Clarity,'' on the other hand, is an apt term for what NHST actually evaluates. \citet{jones2000sensible} and \citet{robinson2001past} suggest that researchers should report $p > 0.05$ using language such as ``the direction of the differences among the treatments was undetermined''. This is a step in the right direction. Replacing ``significance'' with ``clarity'' takes this idea further, and has the promise to substantially improve statistical communication.

@@

@@

\noindent We believe that NHST is useful as a simple, robust way to ask whether an effect can be seen clearly in a particular data set \citep{robinson2001past}, and that careful, clarity-based language can reduce misinterpretation and miscommunication.

@@

\noindent We thank members of the Dushoff and Bolker labs for helpful comments on the first draft of the manuscript.
